---
title: "Assignment 2: Coding Basics"
author: "John Fay"
output: pdf_document
geometry: margin=2.54cm
editor_options: 
  chunk_output_type: console
---

## OVERVIEW

This exercise accompanies the lessons in Environmental Data Analytics on coding basics.

## Directions

1.  Rename this file `<FirstLast>_A02_CodingBasics.Rmd` (replacing `<FirstLast>` with your first and last name).
2.  Change "Student Name" on line 3 (above) with your name.
3.  Work through the steps, **creating code and output** that fulfill each instruction.
4.  Be sure to **answer the questions** in this assignment document.
5.  When you have completed the assignment, **Knit** the text and code into a single PDF file.
6.  After Knitting, submit the completed exercise (PDF file) to Sakai.

## Basics Day 1

1.  Generate a sequence of numbers from one to 100, increasing by fours. Assign this sequence a name.

2.  Compute the mean and median of this sequence.

3.  Ask R to determine whether the mean is greater than the median.

4.  Insert comments in your code to describe what you are doing.

```{r}
#1. Generate sequence...
my_seq <- seq(1,100,4)

#2. Compute mean & median
the_mean <- mean(my_seq)
the_median <- median(my_seq)

#3. Determine equality between mean and median
if (the_mean == the_median){
  print("The mean and median are the same")
} else {
    print("The mean and median are different")
}
```

## Basics Day 2

5.  Create a series of vectors, each with four components, consisting of (a) names of students, (b) test scores out of a total 100 points, and (c) whether or not they have passed the test (TRUE or FALSE) with a passing grade of 50.

6.  Label each vector with a comment on what type of vector it is.

7.  Combine each of the vectors into a data frame. Assign the data frame an informative name.

8.  Label the columns of your data frame with informative titles.

```{r}
#5. Create vectors (6. with labels)
the_names <- c('Aleph','Beta','Gamma','Delta') #chr  (character)
the_scores <- c(34,55,45,66)                   #num  (integer)
the_grade <- c(F,T,F,T)                        #logi (Boolean)

#7. Combine vectors into a dataframe
#Method 1
the_df <- data.frame(the_names, the_scores, the_grade)
#Method 2
the_df2 <- as.data.frame(cbind(the_names, the_scores, the_grade))

#8. Label the columns
names(the_df) = c("student","score","grade")
```

9.  QUESTION: How is this data frame different from a matrix?

> Answer:

10. Create a function with an if/else statement. Your function should take a **vector** of test scores and print (not return) whether a given test score is a passing grade of 50 or above (TRUE or FALSE). You will need to choose either the `if` and `else` statements or the `ifelse` statement. 

11. Apply your function to the vector with test scores that you created in number 5.

```{r}
#10. Create a function to compute a vector of P/F from a vector of scores
compute_passfail <- function(scores){
  the_result <- ifelse(scores > 50,"Pass","Fail")
  print(the_result)
}

#11. Apply the function
compute_passfail(the_df$score)
# -or-
ifelse(the_df$score > 50,"Pass","Fail")
```

12. QUESTION: Which option of `if` and `else` vs. `ifelse` worked? Why?

> Answer:
